---
title: "Cloud Computing Final (Exploratory Analysis)"
author: "Connor Smith"
date: "2023-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data}
day.data = read.csv('day.csv', header=TRUE)
library(ggplot2)
library(MASS)
attach(day.data)
```

## Normality Tests

season, yr, mnth, holiday, weekday, workingday, and weathersit are all categorical variables (encoded), so they should not be transformed.

temp, atemp, hum, and windspeed should all be assessed.

```{r normality check for windspeed}
# Check normality
qqnorm(day.data$windspeed)
qqline(day.data$windspeed)

# also trying the shapiro normality test
shapiro.test(day.data$windspeed)

# Transformation check
boxcox_result <- boxcox(day.data$windspeed ~ 1)
# optimal lambda
boxcox_result$x[which.max(boxcox_result$y)]

# transformation
day.data$tf_windspeed <- (day.data$windspeed^0.424 - 1) / 0.424


# checking transformed normality
qqnorm(day.data$tf_windspeed)
qqline(day.data$tf_windspeed)
shapiro.test(day.data$tf_windspeed)
```


```{r normality check for hum}
# Check normality
qqnorm(day.data$hum)
qqline(day.data$hum)

# also trying the shapiro normality test
shapiro.test(day.data$hum)

# Transformation check
boxcox_result <- boxcox(day.data$hum ~ 1)
# optimal lambda
boxcox_result$x[which.max(boxcox_result$y)]
```


```{r model without interaction}
no_interaction.glr <- lm(casual ~ tf_windspeed + holiday + weekday + workingday + weathersit + hum + temp, data=day.data)
summary(no_interaction.glr)
```


```{r individual correlations}

cor(day.data$casual, day.data$tf_windspeed)
no_interaction.glr <- lm(casual ~ tf_windspeed, data=day.data)
summary(no_interaction.glr)

cor(day.data$casual, day.data$holiday)
no_interaction.glr <- lm(casual ~ holiday, data=day.data)
summary(no_interaction.glr)

cor(day.data$casual, day.data$weekday)
no_interaction.glr <- lm(casual ~ weekday, data=day.data)
summary(no_interaction.glr)

cor(day.data$casual, day.data$workingday)
no_interaction.glr <- lm(casual ~ workingday, data=day.data)
summary(no_interaction.glr)

cor(day.data$casual, day.data$weathersit)
no_interaction.glr <- lm(casual ~ weathersit, data=day.data)
summary(no_interaction.glr)

cor(day.data$casual, day.data$hum)
no_interaction.glr <- lm(casual ~ hum, data=day.data)
summary(no_interaction.glr)

cor(day.data$casual, day.data$temp)
no_interaction.glr <- lm(casual ~ temp, data=day.data)
summary(no_interaction.glr)
```


```{r model without interaction}
no_interaction.glr <- lm(casual ~ tf_windspeed + workingday + weathersit + temp, data=day.data)
summary(no_interaction.glr)
```

```{r boxplots for day of the week}
# Create a boxplot
boxplot(casual ~ weekday, data = day.data,
        main = "Boxplot of Casual vs. Weekday",
        xlab = "Weekday",
        ylab = "Casual",
        col = "blue")
```
It looks like there is a dependency here, but the regression may not pick up on it since the highest and lowest number correlates with the increased rentals. We could recode these so that 0 and 6 (weekend days) = 0 and 1-5 (weekdays) = 1, however that seems to be the point of the 'workingday' column:


```{r boxplots for working day}
# Create a boxplot
boxplot(casual ~ workingday, data = day.data,
        main = "Boxplot of Casual vs. Workingday",
        xlab = "Workingday",
        ylab = "Casual",
        col = "blue")
```

This is the exact same data, just recoded. This would be better in our regression and weekday might be best left out.